---
title: "test fingertips markdown"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(fingertipsR)
library(fingertipscharts)
library(dplyr)
library(plotly)
library(glue)
library(httr)

wanted_indicators <- c(93085, 92517, 93014, 90366, 90282, 22401, 40401, 30314)

# Get data for Bolton, GM combined authority and England 
fingertips_data <- fingertips_data(IndicatorID = wanted_indicators, ProfileID = 19, AreaTypeID = 102,
                             ParentAreaTypeID = 126, rank = TRUE) %>%
  filter(AreaCode %in% c("E08000001", "E47000001", "E92000001"))

  selected_data <- fingertips_data %>%
    filter(IndicatorID== 22401, Sex == "Persons", Age == "65+ yrs", AreaCode %in% c("E08000001", "E47000001")) %>%
    group_by(AreaName) 
```

## Test markdown document

In the latest year `r selected_data[selected_data$AreaName == "Bolton" & selected_data$TimeperiodSortable == max(selected_data$TimeperiodSortable),"Timeperiod"]`, Bolton had `r selected_data[selected_data$AreaName == "Bolton" & selected_data$TimeperiodSortable == max(selected_data$TimeperiodSortable),"Value"]` emergency hospital admissions due to falls among those aged 65+.  

```{r plotly_chart}
# plotly

# working version  
  pal <- c("red", "green", "black")
  pal <- setNames(pal, c("Worse",  "Better", "NA"))
  myhover <- glue("{selected_data$AreaName}<br>{selected_data$Timeperiod} 
                 {prettyNum(round(selected_data$Value, 1), big.mark = ",")}")  
  
  plot_ly(data = selected_data, x = ~Timeperiod, y = ~Value, 
          color = ~ComparedtoCombinedauthoritiesvalueorpercentiles, colors = pal,
          showlegend = FALSE,
          error_y = ~list(type = "data", symmetric = FALSE,
                          array = c(UpperCI95.0limit-Value),
                          arrayminus = c(Value - LowerCI95.0limit),
                          color = "black",
                          line = list(width = 0.5, color = "red"))
  ) %>%
    add_trace(type='scatter',
              mode='lines+markers',
              marker = list(size = 11, line = list(color = "black", width = 1.5)),
              line = list(color = "black", width = 0.5),
              hoverinfo = "text",
              text = ~myhover
    ) %>%
    layout(title = glue("<b>{selected_data$IndicatorName[1]}</b>"),
           xaxis = list(title = "<b>Time period</b>"),
           yaxis = list(title = "<b>Value</b>"))
```

